### 2021-07-11 src/paper/data/assortativity.tex
Make sections A.1 to A.4 ready for review.

As explained in section \ref{sec:matching}, the probability that two individuals are
matched can depend on background characteristics. In particular, we allow this
probability to depend on age and county of residence. While we do not have good data on
geographical assortativity and just roughly calibrate it such that 80\% of contacts are
within the same county, we can calibrate the assortative mixing by age from
\cite{Mossong2008}.

@@

    \caption{Assortativity by Age Group for Non Recurrent Other and Work Contacts}
    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
        recurrent contacts by age group for other contacts on the left and work contacts
        on the right. A row shows the share of contacts a certain age group has with all
        other age groups. Higher values are colored in darker red tones. The diagonal
        represents the share of contacts with individuals from the same age group. The
        80-100 age group for other contacts was so small that we assumed for them to have
        the same contact distribution as the 70-79 year olds. For work contacts, we only
        show age groups that have a significant fraction of working individuals. The
        matrix must not be symmetric because the groups have very different sizes.}

@@

Our two other types of contacts, households and schools, get their assortativity by
construction. Schools are groups where the same children of the mostly same age and
county meet with teachers every day. Household composition follows directly from the
German microcensus data we use to construct our synthetic population.

>>>>

As explained in section \ref{sec:matching}, the probability that two individuals are
matched can depend on background characteristics. In particular, we allow this
probability to depend on age and county of residence. While we do not have good data on
geographical assortativity and set it such that 80\% of contacts are within the same
county, we can calibrate the assortativity by age from \cite{Mossong2008}.

@@

    \caption{Assortativity by Age Group for Non Recurrent Other and Work Contacts}
    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
        recurrent contacts by age group for other contacts on the left and work contacts
        on the right. A row shows the share of contacts a certain age group has with all
        other age groups. Higher values are colored in darker red tones. The diagonal
        represents the share of contacts with individuals from the same age group. The
        80-100 age group for other contacts was so small that we assumed for them to have
        the same contact distribution as the 70-79 year olds. For work contacts, we only
        show age groups that have a significant fraction of working individuals.}

@@

For recurrent contacts, we constructed groups to have the following features: Recurrent
work contacts are not assortative by age. Daily work groups are always of the same county
and weekly work contacts are to 80\% with workers from the same county. Other recurrent
contacts are constructed the same way but we impose for daily contacts that they are
always with individuals from the same age group. School classes are groups where the same
children of the mostly same age and county meet with teachers every day. Nurseries and
preschools mix children by age but match them to come mostly from the same county.
Household age composition follows directly from the German microcensus data we use to
construct our synthetic population.

### 2021-07-11 src/paper/data/medical_parameters.tex
Make sections A.1 to A.4 ready for review.

This section discusses the medical parameters used in the model, their sources and how we
arrived at the distributions used in the model.\footnotemark See
Figure~\ref{fig:model_disease_progression} for a summary of our disease progression model.

\footnotetext{Additional information can be found in the
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.}

@@

% Latency Period
The first medical parameter we need is the length of the period between infection and the
start of infectiousness, the so called latent period. We infer it from two other measures
that are more common in the medical literature: Firstly, the time between infection and
the onset of symptoms, the incubation period. Secondly, the time between the start of
infectiousness and the onset of symptoms. We assume that the latency period is the same
for symptomatic and asymptomatic individuals.

@@

% Length of the presymptomatic stage
The \href{https://bit.ly/3pi18Ax}{European Centre for Disease Prevention and Control}
reports that people become infectious between one and two days before symptoms
start.\footnotemark

\footnotetext{This is similar to \citet{He2020} and in line with \citet{Peak2020}.}

Taking these estimates together, we arrive at a latent period of one to five days.

@@

We assume that the duration of infectiousness is the same for both symptomatic and
asymptomatic individuals as evidence suggests little differences in the transmission
rates between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral
load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020},
\citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the duration of
infectiousness is based on \citet{Byrne2020}. For symptomatic cases they arrive at zero
to five days before symptom onset (see their figure 2) and three to eight days of
infectiousness afterwards.\footnote{Viral loads may be detected much later but eight days
seems to be the time after which most people are culture negative, as also reported by
\citet{Singanayagam2020}.} Thus, we arrive at 0 to 13 days as the range for
infectiousness among individuals who become symptomatic (see also figure 5).

@@

For the time from symptom onset until need for intensive care we rely on data by the US
CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
%Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of
%requiring intensive care per age group.
%
%\begin{table}[tb]
%    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
%    \label{tab:symptomatic-to-ICU}
%    \centering
%
%    \begin{tabular}{ll}
%        \toprule
%        Age Group & Share \\
%        \midrule
%        0-9 & 0.00005 \\
%        10-19 & 0.00030 \\
%        20-29 & 0.00075 \\
%        30-39 & 0.00345 \\
%        40-49 & 0.01380 \\
%        50-59 & 0.03404 \\
%        60-69 & 0.10138 \\
%        70-79 & 0.16891 \\
%        80-100 & 0.26871 \\
%        \bottomrule
%    \end{tabular}
%
%    \tablenotes{
%        The data is taken from \citet{Stokes2020} and \href{https://bit.ly/354pKTZ}{the
%        OpenABM-Project}. }
%
%\end{table}

For those who will require intensive care we follow \citet{Chen2020} who estimate the
time from symptom onset to ICU admission as 8.5 $\pm$ 4 days. This aligns well with
numbers reported for the time from first symptoms to hospitalization:
\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also
in line with the duration estimates collected by \href{https://bit.ly/3gkGtaU}{the
Robert-Koch-Institut}. We assume that the time between symptom onset and ICU takes 4, 6,
8 or 10 days with equal probabilities. As we do not model nursing homes, do not focus on
matching deaths and do not use the number of individuals in intensive care to estimate
our parameters, these numbers are not important for our empirical results.

@@

We take the survival probabilities and time to death and time until recovery from
intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
match this approximately we discretize that 41\% of individuals who will die from
Covid-19 do so after one day in intensive care, 22\% day after 12 days, 29\% after 20
days and 7\% after 32 days. Again, we rescale this for every age group among those that
will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
days. We discretize this such that of those who recover in intensive care, 22\% do so
after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

>>>>

This section discusses the medical parameters used in the model, their sources and how we
arrived at the distributions used in the model.\footnote{Additional information can be
found in the
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.} See Figure~\ref{fig:model_disease_progression} for a summary of our
disease progression model.

@@

% Latency Period
The first medical parameter we need is the length of the period between infection and the
start of infectiousness, the so called latent period. \cite{Zhao2021} estimate the latent
period to last 3.3 days (95\% CI: 0.2, 7.9). In line with this estimate our latent period
lasts one to five days with a slightly lower average of 2 days.

@@

% Length of the presymptomatic stage
Similar to \citet{Peak2020} and in line with \citet{He2020} we set the length of the
presymptomatic stage to be one to two days.

@@

We assume that the duration of infectiousness is the same for both symptomatic and
asymptomatic individuals as evidence suggests little differences in the transmission
rates between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral
load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020},
\citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the duration of
infectiousness is based on \citet{Byrne2020}. For symptomatic cases they arrive at zero
to five days before symptom onset (see their figure 2) and three to eight days of
infectiousness afterwards.\footnote{Viral loads may be detected much later but eight days
seems to be the time after which most people are culture negative, as also reported by
\citet{Singanayagam2020}.} Thus, we arrive at zero to 13 days as the range for
infectiousness among individuals who become symptomatic (see also figure 5).

@@

For the time from symptom onset until need for intensive care we rely on data by
\cite{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.

For those who will require intensive care we follow \citet{Chen2020} who estimate the
time from symptom onset to ICU admission as $8.5 \pm 4$ days. This aligns well with
numbers reported for the time from first symptoms to hospitalization:
\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
that the time between symptom onset and ICU takes four, six, eight or ten days with equal
probabilities.

@@

We take the survival probabilities and time to death and time until recovery from
intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
match this approximately we discretize that 41\% of individuals who will die from
Covid-19 do so after one day in intensive care, 22\% day after twelve days, 29\% after 20
days and 7\% after 32 days. Again, we rescale this for every age group among those that
will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
days. We discretize this such that of those who recover in intensive care, 22\% do so
after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

\FloatBarrier

### 2021-07-11 src/paper/data/synthetic_population.tex
Make sections A.1 to A.4 ready for review.

We sample households to build our synthetic population of over one million households
keeping for each individual their age, gender, occupation and whether they work on
Saturdays and Sundays. For each household we draw its county and set the corresponding
federal state.% we draw the county because county is not part of the campus file

We randomly assign 35\% of children below three to attend a nursery \citep{Destatis2020}.
For children between three and six years old, we assume all go to preschool (officially
92.5\% according to \cite{Destatis2020}). Nursery groups consist

Children that attend a nursery meet in groups of four plus one adult care taker every
weekday when there are no school vacations. Preschool children meet in groups of nine
with two adult care takers. These groups are mixed with respect to age but all belong to
the same state and mostly to the same county. For nurseries the group size is four, for
preschools it's nine \citep{BertelsmannStiftung2019}.

Every child that goes to school is part of three different classes that meet every day on
weekdays when there are no vacations and no school policies are in place.\footnote{We
implement vacations on the federal state level.} Each class consists of approximately 23
students \citep{OECD2013} and two teachers. All students in a class are of the same age
and live in the same state and if mostly also in the same county. In addition, each
child gets assigned a value that captures his or her need to attend nursery, preschool or
school. This allows us to capture various degrees of emergency care that can be granted
while educational facilities are closed or function in some kind of rotating schedule.

Workers are assigned to a daily meeting work group. The group sizes vary to match the
number of daily repeating work contacts reported by working individuals in
\cite{Mossong2008}. These groups only consist of workers that work in the same county.
For a distribution of the number of daily recurring work contacts see
Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
we match each worker with up to 14 other workers to match the number of reported weekly
work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}. Each pair is
assigned a weekend on which they always meet in the absence of work policies. 80\% of
these contacts are individuals from the same county.
% work contact priority
In the same way children have a educational priority determining if they are entitled to
emergency care workers are assigned a work contact priority that captures how necessary
their work is and to which degree they can work from home. This means that it's always
the same individuals that continue to have work contacts when work from home mandates of
a certain strictness are in place. We use the reduction in work mobility reported by
\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.

In addition to creating groups for educational facilities and work we also have other
recurring contacts to represent things like groups of friends or sports teams that
practice regularly together. Both daily and weekly groups are created analogously to the
work groups but matching the numbers in Figure~\ref{n_contacts_other_daily_recurrent} and
Figure~\ref{n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
highly assortative by age all individuals that have a daily leisure contact are matched
with a person not only from the same county but also from the same age group.

The individuals in our population can react to events such as developing symptoms that
are typical of CoViD-19, a positive PCR test or a positive rapid test by reducing their
contacts. To determine who would reduce their contacts in such a situation or demand a
rapid test to avoid quarantine we introduce a quarantine compliance parameter. Similarly
we introduce a rapid test compliance parameter that determines in which order individuals
start demanding rapid tests when rapid tests become increasingly available. This makes
sure that when for example only 10\% of workers get tested, it's the same workers that
have access to tests every week.

Lastly, for the distribution of vaccinations every individual is assigned a vaccination
group and a vaccination rank from that group that creates a complete vaccination queue
over the population including 15\% that refuse to be vaccinated as reported by the Robert
Koch Institute (RKI, \cite{RKI2021}). The vaccination groups are created to match the
recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
that teachers were prioritized more than recommended by the
commission.}\comment[id=K]{Write how that plays out in our simulations for the different
age groups.}

To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
put adolescents and children into two groups that follow after the general population.
These groups do not become eligible within our simulation frame until June.

>>>>

We sample households to build our synthetic population of over one million households
keeping for each of the 2.3 million individuals their age, gender, occupation and whether
they work on Saturdays and Sundays. For each household we draw its county and set the
corresponding federal state.% we draw the county because county is not part of the campus file

We randomly assign 35\% of children below three to attend a nursery \citep{Destatis2020}.
For children between three and six years old, we assume all go to preschool (officially
92.5\% according to \cite{Destatis2020}).
% nurseries and preschools
Children that attend a nursery meet in groups of four \citep{BertelsmannStiftung2019}
plus one adult care taker every weekday when there are no school vacations. Preschool
children meet in groups of nine \citep{BertelsmannStiftung2019} with two adult care
takers. These groups are mixed with respect to age but all belong to the same state and
mostly to the same county.

% school
Every child that goes to school is part of three different classes that meet every day on
weekdays when there are no vacations and no school policies are in place.\footnote{We
implement vacations on the federal state level.} Each class consists of approximately 23
students \citep{OECD2013} and two teachers. All students in a class are of the same age
and live in the same state and mostly also in the same county. In addition, each
child gets assigned a value that captures his or her need to attend nursery, preschool or
school. This allows us to capture various degrees of emergency care that can be granted
while educational facilities are closed or are on some kind of rotating schedule.

% workers
Workers are assigned to a daily meeting work group. The group sizes vary to match the
number of daily repeating work contacts reported by working individuals in
\cite{Mossong2008}. These groups only consist of workers that work in the same county.
For a distribution of the number of daily recurring work contacts see
Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
we match each worker with up to 14 other workers into pairs to match the number of
reported weekly work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}.
Each pair is assigned a weekday on which they always meet in the absence of work
policies. 80\% of these contacts are individuals from the same county.% work contact priority
In the same way children have an educational priority determining if they are entitled to
emergency care workers are assigned a work contact priority that captures how necessary
their work is and to which degree they can work from home. This means that it's always
the same individuals that continue to have work contacts when work from home mandates of
a certain strictness are in place. We use the reduction in work mobility reported by
\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.

% other recurring contacts
In addition to creating groups for educational facilities and work we also have other
recurring contacts to represent things like groups of friends or sports teams that
practice regularly together. Both daily and weekly groups are created analogously to the
work groups but matching the numbers in Figure~\ref{n_contacts_other_daily_recurrent} and
Figure~\ref{n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
highly assortative by age all individuals that have a daily leisure contact are matched
with a person not only from the same county but also from the same age group.

% individual responses
The individuals in our population can react to events such as developing symptoms that
are typical of CoViD-19, a positive PCR test or a positive rapid test by reducing their
contacts. To determine who would reduce their contacts in such a situation or demand a
rapid test we introduce a quarantine compliance parameter. Similarly we introduce a rapid
test compliance parameter that determines in which order individuals start demanding
rapid tests when rapid tests become increasingly available. This makes sure that when for
example only 10\% of workers get tested, it's the same workers that have access to tests
every week.

Lastly, for the distribution of vaccinations every individual is assigned a vaccination
group and a vaccination rank from that group that creates a complete vaccination queue
over the population including 15\% that refuse to be vaccinated as reported by the Robert
Koch Institute (RKI, \cite{RKI2021}). The vaccination groups are created to match the
recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
that teachers were prioritized more than recommended by the commission.} To cover that
the Pfizer-BioNTech vaccine was later approved for younger age groups we put adolescents
and children into two groups that follow after the general population. These groups do
not become eligible within our simulation frame until June. The way vaccinations are rolled out in our model is shown in Figure~\ref{fig:vaccinations_by_age_group}.
